Bayes Function Testing

### 2021-06-18 Bayes Function Testing.Rmd
testing bayes function on 4 samples created with extreme proportions of signatures 18 and 21
Updated testing with finer range of exposures and expected output for each chunk.

```{r}
create_test_sample <- function(signatures, prop, num) {
  cosmic_signatures <- load_cosmic_matrix()
  process <- cosmic_signatures[, signatures]

@@

```{r}
#Create a set of 4 test samples with extreme propotions of COSMIC signatures 18 and 21
test_sample <- create_test_sample(c(18,21), c(1,0,0,1,.9,.1,.1,.9),4)
test_sample
```

@@

#set column and row names to names of signatures
colnames(test_nmf$signatures) <- c("Signature 21", "Signature 18")
rownames(test_nmf$contribution) <- c("Signature 21", "Signature 18")
```

@@

```{r}
#G[C>A]A mutation is the most common in Signature 18
#Sample 1 is ~100% Signature 18
extract_all_prob("G[C>A]A", 1, test_nmf)
#Sample 3 is ~90% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
```

```{r}
#G[T>C]A mutation is the most common in Signature 21
#Sample 2 is ~100% Signature 21
extract_all_prob("G[T>C]A", 2, test_nmf)
#Sample 4 is ~90% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
```

>>>>

```{r}
#Creates samples/mutational profiles using selected signatures from COSMIC
#and desired contributions from each signature.
create_test_sample <- function(signatures, prop, num) {
  cosmic_signatures <- load_cosmic_matrix()
  process <- cosmic_signatures[, signatures]

@@

```{r}
#Create a set of 7 test samples with extreme proportions of COSMIC signatures 18 and 21
test_sample <- create_test_sample(c(18,21), c(  1,  0,
                                              0.9,0.1,
                                              0.7,0.3,
                                              0.5,0.5,
                                              0.3,0.7,
                                              0.1,0.9,
                                                0,  1),
                                  num = 7)
test_sample
```

@@

#set column and row names to names of signatures
colnames(test_nmf$signatures) <- c("Signature 18", "Signature 21")
rownames(test_nmf$contribution) <- c("Signature 18", "Signature 21")
```

@@

## Proportions should lean in favor for signature 18 and decrease with its sample size

```{r}
#G[C>A]A mutation is the most common in Signature 18
#1287 times more likely in signature 18 than 21
#Sample 1 is ~100% Signature 18
extract_all_prob("G[C>A]A", 1, test_nmf)
#Sample 2 is ~90% Signature 18
extract_all_prob("G[C>A]A", 2, test_nmf)
#Sample 3 is ~70% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
#Sample 3 is ~50% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
#Sample 4 is ~30% Signature 18
extract_all_prob("G[C>A]A", 4, test_nmf)
#Sample 5 is ~10% Signature 18
extract_all_prob("G[C>A]A", 5, test_nmf)
#Sample 6 is ~0% Signature 18
extract_all_prob("G[C>A]A", 6, test_nmf)
```

## Proportions should lean in favor for signature 21 and decrease with its sample size

```{r}
#G[T>C]A mutation is the most common in Signature 21
#269 times more likely in signature 21 than 18
#Sample 6 is ~100% Signature 21
extract_all_prob("G[T>C]A", 6, test_nmf)
#Sample 5 is ~90% Signature 21
extract_all_prob("G[T>C]A", 5, test_nmf)
#Sample 4 is ~70% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
#Sample 4 is ~50% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
#Sample 3 is ~30% Signature 21
extract_all_prob("G[T>C]A", 3, test_nmf)
#Sample 2 is ~10% Signature 21
extract_all_prob("G[T>C]A", 2, test_nmf)
#Sample 1 is ~0% Signature 21
extract_all_prob("G[T>C]A", 1, test_nmf)
```

```{r}
#Find smallest difference between contribution rate for every mutation type between
#signature 18 and 21. Attempting to find mutation type w/ equal probabilities for both.
cosmic_signatures <- load_cosmic_matrix()
(process18 <- cosmic_signatures[, c(18)])
(process21 <- cosmic_signatures[, c(21)])
diff_signatures <- abs(process21-process18)
diff_signatures[93]=1 #contribution for signature 21 is 0 (T[T>G]A)
diff_signatures[85]=1 #contribution for signature 21 is 0
which.min(diff_signatures)
```

## Signature 21 probabilities should be zero for all samples

```{r}
#LOW probability of T[T>G]A mutation in signature 18
#zero probability of T[T>G]A mutation in signature 21
#100% Signature 18 and 0% Signature 21
extract_all_prob("T[T>G]A", 1, test_nmf)
#90% Signature 18 and 10% Signature 21
extract_all_prob("T[T>G]A", 2, test_nmf)
#70% Signature 18 and 30% Signature 21
extract_all_prob("T[T>G]A", 3, test_nmf)
#50% Signature 18 and 50% Signature 21
extract_all_prob("T[T>G]A", 4, test_nmf)
#30% Signature 18 and 70% Signature 21
extract_all_prob("T[T>G]A", 5, test_nmf)
#10% Signature 18 and 90% Signature 21
extract_all_prob("T[T>G]A", 6, test_nmf)
#0% Signature 18 and 100% Signature 21
extract_all_prob("T[T>G]A", 7, test_nmf) #BAD RESULT
```

## Probabilities should almost reflect contribution proportions

```{r}
#Low chances of C[C>G]A mutations in both but roughly equal chances of mutations
#(4.58 more likely in signature 18, as opposed to hundreds or thousands times more likely)
#100% Signature 18 and 0% Signature 21
extract_all_prob("C[C>G]A", 1, test_nmf)
#90% Signature 18 and 10% Signature 21
extract_all_prob("C[C>G]A", 2, test_nmf)
#70% Signature 18 and 30% Signature 21
extract_all_prob("C[C>G]A", 3, test_nmf)
#50% Signature 18 and 50% Signature 21
extract_all_prob("C[C>G]A", 4, test_nmf)
#30% Signature 18 and 70% Signature 21
extract_all_prob("C[C>G]A", 5, test_nmf)
#10% Signature 18 and 90% Signature 21
extract_all_prob("C[C>G]A", 6, test_nmf)
#0% Signature 18 and 100% Signature 21
extract_all_prob("C[C>G]A", 7, test_nmf)
```
